---
title: "My functions in foofactors"
author: "Sukey"
output:
  github_document:
    toc: yes
---

# Installtion
```{r}
library(dplyr)
library(devtools)
# devtools::install_github("STAT545-UBC-students/hw07-Sukeysun")
```

# Functions in foofactor
Factors are a very useful type of variable in R, but they can also drive you nuts. Especially the "stealth factor" that you think of as character.

Can we soften some of their sharp edges?

## fbind
Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work.  Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

## freq_out

The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```


## detect_factor

The `detect_factor` function is a function to test whether the factor is character.

```{r}
x <- c('a', 'b', 'b')
x_fact <- factor(x)
y <- c('a', 'b')
y_fact <- factor(y)
detect_factor(x_fact)
detect_factor(y_fact)
```
x_fact is  factor while y_fact is character as the length of unique values of y_fact is equal to the length of y_fact

## reorder_factor

The `reorder_factor` returns the input factor in descending order
```{r}
x <- c('c', 'b', 'd')
x_fact <- factor(x)
y <- c('c','c','d')
y_fact <- factor(y)
levels(y) <- c('d','c')
print("The original order of factor")
levels(x_fact)
levels(y_fact)
print("Reorder the input factor")
reorder_factor(x_fact)
reorder_factor(y_fact)
  
```

## set_factor
This function  sets levels to the order in which they appear in the data, i.e. set the levels “as is”
```{r}
x <- c("b","c","a")
x_fact <- factor(x)
print("The original order of factor")
levels(x_fact)
print("Set factor")
set_factor(x_fact)
```

The original order of x levles is set in an increasing order. Now the it is set as the sequence that each level appears.

## read_dataframe
read data frames to plain text delimited files while retaining factor levels

```{r}
gdURL <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
df <- read_dataframe(gdURL)


data.frame( variable = names( df ),
           classes = sapply( df, class ),
           factorlevel = sapply( df, nlevels ),
           first_values = sapply( df, function( x ) paste0( head( x ),  collapse = ", ") ),
           row.names = NULL ) %>%
  knitr::kable()
```

We can see that the country and continent is read as factor.

## write_dataframe
write data frames to plain text delimited files while retaining factor levels. The dataframe will be named as "written_dataframe.txt"
```{r}
df <-  data.frame(x = (c("a","b","c")),y=c("d","e","f"))
write_dataframe(df)
```

